# Syllabus {#syllabus}

<!--
You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter 

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].
-->

Here you will find the schedule and the topics that will be covered in each lesson.


## Schedule

### May {-}

| Monday 	| Tuesday 	| Wednesday 	| Thursday 	| Friday 	| Saturday 	| Sunday 	|
|--------	|---------	|-----------	|----------	|--------	|----------	|--------	|
|        	|         	|           	|          	|        	|          	|        	|
|        	|         	|           	|          	|        	|          	|        	|
|        	|         	|           	|          	|        	|          	|        	|
| 21     	| 22      	| 23        	| 24       	| 25     	| 26       	| 27     	|
| 28     	| 29      	| 30        	| 31       	|        	|          	|        	|


### June {-}

| Monday 	| Tuesday 	| Wednesday 	| Thursday 	| Friday 	| Saturday 	| Sunday 	|
|--------	|---------	|-----------	|----------	|--------	|----------	|--------	|
|        	|         	|           	|          	| 1      	| 2        	| 3      	|
| 4      	| 5       	| 6         	| 7        	| 8      	| 9        	| 10     	|
| 11     	| 12      	| 13        	| 14       	| 15     	| 16       	| 17     	|
| 18     	| 19      	| 20        	| 21       	| 22     	| 23       	| 24     	|
| 25     	| 26      	| 27        	| 28       	| 29     	| 30       	|        	|

### July {-}

### August {-}

## Lesson Breaksown

### Lesson 1

- Getting setup
  - Getting the shell
  - SSH keys
- The SDAL Infrastructure
  - RStudio
  - The other tools availiable
- Shell (Bash)
- The "Tidyverse"
- Exploring your data through visualizations (ggplot2)
- Loading a dataset (readr)

### Lesson 2

- Markdown
- Knitr
- The project template
- Projects
- Git (locally)

### Lesson 3

- Git (remotes)
- Git (branches)
- Git (collaboration)

### Lesson 4

- Workflow Basics
- Transformations
- Scripts

### Lesson 5

- EDA
- Tibbles
- Tidy

### Lesson 6

- Relational Data
- SQL

### Lesson 7

- Strings
- Factors
- Dates and Times

### Lesson 8

- Pipes
- Functions
- Vectors
- Iteration

### Lesson 9

- Communication
  - R Markdown
  - Graphics
- R Markdown formats
- Shiny

### Lesson 10
