# Data Transformations

Data Transformation chapter in r4ds

- http://r4ds.had.co.nz/transform.html

DataCamp Courses:

- https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial
- https://www.datacamp.com/courses/introduction-to-the-tidyverse
- https://www.datacamp.com/courses/cleaning-data-in-r

References:

- https://dplyr.tidyverse.org/
- [Rstudio Data Transformation Cheat Sheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
- [Tidyverse for beginners DataCamp Cheatsheet](http://datacamp-community.s3.amazonaws.com/e63a8f6b-2aa3-4006-89e0-badc294b179c)
- 

## `dplyr` library

```{r}
library(dplyr)
library(nycflights13)
```

```{r}
flights
```

The basic verbs in `dplyr`

- `filter()`: selects rows
- `arrange()`: reorders rows
- `select()`: selects (and re-order) columns
- `mutate()`: create new variables (columns) based on existing variables (columns)
- `summarise()`: collapses multiple values into a single value (e.g., mean, standard deviation, etc)

## Filter

```{r}
filter(flights, month == 1)
```

```{r}
filter(flights, month == 1, day == 1)
```


```{r}
jan1 <- filter(flights, month == 1, day == 1)
# jan1 = filter(flights, month == 1, day == 1)
```


```{r}
(jan1 <- filter(flights, month == 1, day == 1))
```

## comparisons

```
> >= < <= != ==
```


```{r}
sqrt(2)^2 == 2
```


```{r}
near(sqrt(2)^2, 2)
```

## logical operators

```
| &
```

```{r}
filter(flights, month == 11 | month == 12)

# filter(flights, month == 11 | 12)  ## this is wrong and will not work like you expect
```


```{r}
filter(flights, month == 11 | 12)
```


```{r}
filter(flights, month %in% c(11, 12))
```

```{r}
filter(flights, arr_delay <= 120, dep_delay <= 12)
```

by default filter will also drop missing values.
See the r4ds chapter for this.

## Arrange

```{r}
arrange(flights, year, month, day)
```


```{r}
arrange(flights, year, month, desc(day))
```

```{r}
arrange(flights, year, desc(month), day)
```

## Select

```{r}
select(flights, year, month, day)
```


```{r}
select(flights, year:day, arr_delay)
```

```{r}
select(flights, -year)
```

```{r}
select(flights, -(year:day))
```

Other functions you can use within select:

```
starts_with
ends_with
contains
matches
num_range # for example to create x1, x2, x3
```

```{r}
rename(flights, "tail_num" = tailnum, 'y' = year)
```

```{r}
select(flights, time_hour, air_time, everything())
```

## Mutate

```{r}
(flights_sml <- select(flights,
                      year:day,
                      ends_with('delay'),
                      distance,
                      air_time))
```

```{r}
mutate(flights_sml,
       gain = arr_delay - dep_delay,
       speed = distance / air_time * 60)
```

```{r}
mutate(flights_sml,
       gain = arr_delay - dep_delay,
       speed = distance / air_time * 60,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )
```

## Summarize (summarise)

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

## Groupby

```{r}
by_day <- group_by(flights,
                   year, month, day)
```

```{r}
summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

Do a group by and perform multiple summarizations

```{r}
by_month <- group_by(flights,
                   year, month)
by_month <- summarize(by_month,
          delay = mean(dep_delay, na.rm = TRUE),
          delay_std = sd(dep_delay, na.rm = TRUE)
          )
by_month
```

The above code can be re-written using the pipe, `%>%`

```{r}
# i'm pretty sure this is easier to read and understand
by_month <- group_by(flights,
                     year, month) %>%
    summarize(delay = mean(dep_delay, na.rm = TRUE),
              delay_std = sd(dep_delay, na.rm = TRUE))
by_month
```

Otherwise you will have to create a temp variable,
or write a nested expression

```{r}
summarize(group_by(flights, year, month),
          delay = mean(dep_delay, na.rm = TRUE),
          delay_std = sd(dep_delay, na.rm = TRUE))
```
