Resources:
More detailed phone number example: https://stackoverflow.com/questions/16699007/regular-expression-to-match-standard-10-digit-phone-number
Excellent cheat sheet: https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf
More than you ever wanted: https://www.regular-expressions.info/tutorial.html

A great package for regex in r is Hadley Wickham's 'stringr' package. We'll be using this often.
```{r}
library(stringr)
```

The main idea behind regular expressions (or regex) is to find blocks of text that match a certain pattern. This pattern can be as simple as a single letter or number, or as complicated as an entire address.

allchars = '1234567890abcdefghijklmnopqrstuvwxyz`,./!@#$%^&*()_+][}{|'
exampleText = "Some phone numbers are 7608675309, perhaps also 403-596-4038. Some people use periods in their numbers, so 402.367.5039. We can also have things like 304 385 1029 or (760)-581-3957. How can we extract all of these?"
```









## Regex in the wild!
Perhaps you've looked at some of Dan's old gitignore files?
```{r}
readLines('./data/regexClips/exampleGitignore')
```

Regex is great for getting a list of files from a directory. For instance, if you want to write a loop to read in all .csv files.
```{r}
list.files(pattern = ".Rmd")
```
